---
title: "Homework 2"
subtitle: "Autumn 2022"
author: "Tanner Huck"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

* * * 

### Instructions

- This homework is due in Gradescope on Wednesday Oct 19 by midnight PST.

- Please answer the following questions in the order in which they are posed. 

- Don't forget to knit the document frequently to make sure there are no compilation errors. 

- When you are done, download the PDF file as instructed in section and submit it in Gradescope. 


* * *

### Exercises 

1. Suppose 12 coins are tossed and the outcome (head or tail) is recorded for each.  
   
a. The sample space $S$ corresponding to this "experiment" consists of all possible sequences of heads and tails that result from tossing 12 coins. How many elements are in $S$? Calculate this number in a code chunk and also explain your answer very briefly.

```{r coin_flip_sample_space}
total_elements = 2^12
total_elements
```

Using the equally likely rule, we know that flipping a coin will result in 2 outcomes with equal probability. Flipping a coin 12 times in a row will simply result in $2_1*2_2*2_3*\dots*2_{12}$ or $2^{12}$ different outcomes with equal probability. Meaning that the sample space will have `r total_elements` elements.

   
b. Let $E$ denote the event that 7 of the 12 coins land on heads. How many elements are in $E$? Explain your answer very briefly.
    
```{r event_seven_heads}
element_seven_heads = choose(12, 7)
element_seven_heads
```

If E is the event that 7 of the 12 coins land of heads, than the number of elements in event E is 12 choose 7. Meaning there are `r element_seven_heads` elements in event E.

c. Calculate $P(E)$ assuming all the elements in $S$ are equally likely. Report your final answer in a sentence using inline code.

```{r proability_element_event_E}
probability_E = element_seven_heads / total_elements
probability_E
```

The probability of event E is `r probability_E`.
   
   
2. To estimate the number $N$ of goldfish in a pond, $r=25$ fish were caught, tagged and released. Later, a second sample of $n=20$ fish were caught and 5 fish in this sample were noted to be tagged.  
 
a.  How many possible samples of size $n=20$ can be formed from the $N$ fish in the pond?  (Leave your answer in terms of a binomial coefficient - you cannot calculate it because you don't know $N$)

To see how many samples of size $n=20$ can be formed from $N$ fish, we simply do $N \choose 20$.
    
b. The event $E$ contains all the samples which have 5 tagged and 15 untagged fish. How many elements are in the event $E$? (Leave your answer in terms of $N$)

Let event E be the event that the sample consists of 5 tagged and 15 untagged fish. Than to find the number of elements in E we can do ${25 \choose {5}}{{N-25} \choose 15}$.
   
c. Assuming each possible sample is equally likely, give an expression for $P(E)$. (Leave your answer in terms of $N$)

Assuming that each sample is equally likely, than the probability of E would be  
$P(E)= {25 \choose 5} {{N-25} \choose 15} / {N \choose 20}$



d.  In this part, we will examine visually how $P(E)$ varies as a function of $N$.  Fill in the blanks in the R code provided and run it in R to create the plot. You should remove the `eval = FALSE` chunk option before knitting.   (Note: this part does not require you to know **tidyverse** or **ggplot** functions, and should be doable by all.)


```{r fish_plot}

fishes <- tibble( #enhanced data frame
        N = 40:200,    #possible values for N: 40,41, ...,200
        prob = choose(25,5)*choose((N-25),15)/choose(N,20))  #write expression for P(E) in terms of N

ggplot(data = fishes, 
       mapping = aes(x = N,    
                     y = prob) ) +
           geom_line() +
             labs( title = "Probability of Fish Sample" ,
                  x = "Number of fish in Pond", 
                  y = "Probability 5 tagged and 15 untagged fish") +
          theme(plot.title=element_text(hjust=0.5))
              
```


3. Among all students seeking treatment at Hall Health, 0.5% are eventually diagnosed as having mononucleosis (event $A$). Of those who do have mono, 90% complain of a sore throat (event $B$). But 30% of those not having mono also have sore throats. 

a. Make a tree diagram of the probabilities relating presenting with a sore throat and a diagnosis of mononucleosis. (Don't forget to include the **openintro** package in the setup chunk)

```{r treediagram_health}
library(openintro) 
treeDiag(main = c("Mononucleosis Status", "Throat"),
         p1 = c(0.005, 0.995),
         p2 = list(c("0.90", "0.10"), c("0.30", "0.70")),
         out1 = c("Present", "Absent"),
         out2 = c("Sore", "Not Sore"),
         showSol = FALSE,
         showWork = TRUE,
         solwd = 0.5,
         digits = 6)
```
    
b. If a student comes to Hall Health and says that she has a sore throat, what is the probability that she has mono? Be sure to show your steps carefully. 

We want to find the conditional probability that someone has mono given that they have a sore throat denoted by $P(A|B)$. By dediniton of conditional probability we can rewrite this as $P(A \cap B)/P(B)$. Now solving for this probability.

```{r probability of having mono given you have a sore throat}
p_A_and_B = 0.005 * 0.90 #Probability of mono and sore throat
p_Ac_and_B = 0.995 * 0.30 #Probability of no mono and sore throat
p_B = p_A_and_B + p_Ac_and_B #Probability of sore throat
p_A_given_B = p_A_and_B / p_B #Probability of mono given that you have sore throat
p_A_given_B
```

Hence the probability that you have mono given you have a sore throat is `r p_A_given_B`.

4. Prove that if $A$ and $B$ are independent events, then the following pairs are also independent. 

a. $A$ and $B^c$

By definition, we can write $A$ as the union of two disjoint sets: $$A = (A \cap B) \cup (A \cap B^c)$$ By using the third axiom of probability and because we know $(A \cap B)$ and $(A \cap B^c)$ are disjoint, we can rewrite as $$P(A) = P(A \cap B) + P(A \cap B^c)$$ Because $A$ and $B$ independent, by the definition of independence we can rewrite $P(A \cap B)$ as $P(A) * P(B)$, which gives us $$P(A) = P(A) * P(B) + P(A \cap B^c)$$ If we isolate $P(A \cap B^c)$ and rearrange we get $$P(A \cap B^c) = P(A) - P(A) * P(B)$$. Then factoring a $P(A)$ gives us $$P(A \cap B^c) = P(A) * (1-P(B))$$ finally by the rule of complements we can rewrite as $$P(A \cap B^c) = P(A) * P(B)$$ Hence we know that $P(A \cap B^c)$ is equal to $P(A) * P(B^c)$. By our definition of independence that means that $A$ and $B^c$ are independent.

b. $A^c$ and $B$

This case is very similar to problem a, so I will just show the steps. (Just switch A and B).
$$B = (B \cap A) \cup (B \cap A^c)$$
$$P(B) = P(B \cap A) + P(B \cap A^c)$$
$$P(B) = P(B) * P(A) + P(B \cap A^c)$$
$$P(B \cap A^c) = P(B) - P(B) * P(A)$$
$$P(B \cap A^c) = P(B) * (1-P(A))$$
$$P(B \cap A^c) = P(B) * P(A^c)$$
Hence by our definition of independence we know that $B$ and $A^c$ are independent as well.

c. $A^c$ and $B^c$.

Using DeMorgan's laws we know that 
$$P(A^c \cap B^c) = P((A \cup B)^c)$$
Then using the complement rule we know 
$$P(A^c \cap B^c) = 1 - P(A \cup B)$$
Using the addition rule
$$P(A^c \cap B^c) = 1 - [P(A) + P(B) - P(A \cap B)]$$
Because we know that $A$ and $B$ are independent we can rewrite as
$$P(A^c \cap B^c) = 1 - [P(A) + P(B) - P(A) * P(B)]$$
Then using the complement rule again, we can substitute all $P(A)$ with $1-P(A^c)$ and likewise for $B$
$$P(A^c \cap B^c) = 1 - [1-P(A^c) + 1-P(B^c) - [(1-P(A^c)) * (1-P(B^c))]$$
Now if we expand the multiplication and then simplify we get
$$P(A^c \cap B^c) = 1 - (1 - P(A^c) * P(B^c))$$
Which is the same as 
$$P(A^c \cap B^c) = P(A^c) * P(B^c)$$
Thus, by our definition of independence, this means that $A^c$ and $B^c$ must be independent.

